---
title: "COVID-19 data from CANSIM Table 13-10-0766-01"
author: "Joel Barnes"
date: '`r format(Sys.time(), "%B %d, %Y") `'
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 5
    code_folding: show
editor_options:
  chunk_output_type: console
---

<style>
  html, body { font-size: 14px; }
  p { margin-bottom: 18px; }
  h1 { margin-top: 30px; font-size: 1.75rem; }
  h1.title {font-size: 2rem;  }
  h2 { font-size: 1.35rem; margin-top: 30px; }
  h3 { font-size: 1.15rem; margin-top: 30px; }
  h4 { font-size: 1.05rem; margin-top: 30px; }
  h4.date { margin-top: 0; margin-bottom: 3rem; }
  p.caption { margin-top: 2rem; }
  img { width: 100%; height: auto; }
  blockquote { border-left: 5px solid #337ab7; border-bottom: 1px solid #efefef; font-weight: normal; font-size: 1rem; }
  .dataTables_filter, .dataTables_length {font-weight: normal; }
  div.main-container { width: 100% !important; max-width: 100% !important; margin-left: auto; margin-right: auto; }
  td.dt-top { vertical-align: top; }
  .level3 h3, .level4 h4 { color: #337ab7; font-weight: bold; }
  .level4 h4 { margin-top: 96px;}
  li > img { margin-top: 24px; margin-bottom: 24px; }
</style>

<img src="https://www.statcan.gc.ca/wet-boew4b/assets/sig-blk-en.svg" style = "max-width: 300px; height: auto; margin-bottom: 24px;" alt="logo">

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dpi = 300)
```

# Load packages to extend base R

```{r}
# Define packages that will be used to extend base R
package_names <- c("cansim", "DT", "ggplot2", "plyr", "plotly", "scales", "tidyverse", "xlsx") 

# Install any packages that do not exist
install_packages <- lapply(package_names, FUN = function(x) if(! require(x, character.only = TRUE)) install.packages(x))

# Load the packages
load_packages <- lapply(package_names, require, character.only = TRUE)
```

# Import CANSIM table

```{r}
d <- get_cansim("13-10-0766-01", refresh = TRUE)
```

# Data wrangling
## Reshape data from long to wide format

```{r}
d_wide <- spread(d %>% select("Case identifier number", "Case information", VALUE, REF_DATE), "Case information", VALUE)
```

## Add leading zeros to case identifier number

```{r}
d_wide$`Case identifier number` <- str_pad(d_wide$`Case identifier number`, width = nchar(max(as.numeric(d$`Case identifier number`))), pad = "0")
```

## Restructure select vectors as factors

```{r}
# Identify vectors
vectors_to_factor <- c("Age group", "Gender", "Transmission", "Hospitalization", "Intensive care unit", "Death")

# Restructure as factors
d_wide[vectors_to_factor] <- lapply(d_wide[vectors_to_factor], factor)

# Add semantic labels
d_wide$`Age group` <- revalue(d_wide$`Age group`, c("1" = "0-19", "2" = "20-29", "3" = "30-39", "4" = "40-49", "5" = "50-59", "6" = "60-69", "7" = "70-79", "8" = "80+", "99" = "Not stated"))
d_wide$Gender <- revalue(d_wide$Gender, c("1" = "Male", "2" = "Female", "7" = "Non-binary", "9" = "Not stated"))
d_wide$Transmission <- revalue(d_wide$Transmission, c("1" = "Travel exposure", "2" = "Community exposure", "3" = "Pending"))
d_wide$Hospitalization <- revalue(d_wide$Hospitalization, c("1" = "Yes", "2" = "No", "9" = "Not stated"))
d_wide$`Intensive care unit` <- revalue(d_wide$`Intensive care unit`, c("1" = "Yes", "2" = "No", "9" = "Not stated"))
d_wide$Death <- revalue(d_wide$Death, c("1" = "Yes", "2" = "No", "9" = "Not stated"))
```

## Create episode date vector

```{r}
# Add day, month and reference year vectors together and structure as a date object
d_wide$`Episode date` <- as.Date(paste0(d_wide$REF_DATE, "-", str_pad(d_wide$`Episode date - month`, 2, pad = "0"), "-", str_pad(d_wide$`Episode date - day`, 2, pad = "0")), format = "%Y-%m-%d")

# Change format to %d-%b-%y
d_wide$`Episode date` <- format(d_wide$`Episode date`, format = "%d-%b-%y")
```

## Remove unwanted vectors from data

```{r}
d_wide <- d_wide %>% select("Case identifier number", "Episode date", Gender, "Age group", Transmission, Hospitalization, "Intensive care unit", Death)
```

## Rename vectors

```{r}
names(d_wide) <- c("CaseID", "Episode Date", "Gender", "Age Group", "Exposure Setting", "Hospitalized", "Intensive Care Unit", "Death")
```

## Order data by case ids in ascending order

```{r}
d_wide <- d_wide %>% arrange(CaseID)
```

## Export data to Excel

```{r}
write.xlsx2(as.data.frame(d_wide), paste0("c:/users/joel/google drive/github/covid19/Table 13-10-0766-01 - updated ", format(Sys.time(), "%Y-%m-%d"), ".xlsx"), row.names = FALSE, showNA = FALSE)
```

# Sortable/searchable raw data table

```{r}
# Output data to JavaScript datatable
datatable(d_wide, 
  extensions = c("Buttons", "Scroller"), 
  options = list(
    pageLength = 25, 
    dom = "Bfrtip", 
    buttons = c("colvis", "copy", "csv", "excel", "pdf"), 
    deferRender = TRUE, 
    searchDelay = 500,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#fff', 'color': '#111'});",
      "}"
    ),
    columnDefs = list(
      list(visible = FALSE, targets = c())
    )
  ), 
  rownames = FALSE,
  escape = FALSE,
  editable = TRUE
)
```

# Incidence
## Sortable/searchable data table by age group

```{r}
# Convert episode date to date object
d_wide$`Episode Date` <- as.Date(d_wide$`Episode Date`, format = "%d-%b-%y")

# Remove cases with no episode date or with an age group value of "Not stated
d_wide <- d_wide %>% filter(! is.na(`Episode Date`) & `Age Group` != "Not stated")

# Drop "Not stated" level from the age group factor
d_wide$`Age Group` <- droplevels(d_wide$`Age Group`, "Not stated")

# Sort data by episode date
d_wide <- d_wide[order(d_wide$`Episode Date`),]

# Collapse several age group levels
#d_wide$`Age Group` <- fct_collapse(d_wide$`Age Group`, "20-59" = c("20-29", "30-39", "40-49", "50-59"))

# Create a crosstab
crosstab <- d_wide %>% group_by(`Age Group`, `Episode Date`) %>% tally()

# Rename the n vector
names(crosstab)[ncol(crosstab)] <- "Incidence"

# Compute cumulative incidence
crosstab <- crosstab %>% group_by(`Age Group`) %>% mutate(`Cumulative Incidence` = cumsum(Incidence))

# Output data to JavaScript datatable
datatable(crosstab, 
  extensions = c("Buttons", "Scroller"), 
  options = list(
    pageLength = 25, 
    dom = "Bfrtip", 
    buttons = c("colvis", "copy", "csv", "excel", "pdf"), 
    deferRender = TRUE, 
    searchDelay = 500,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#fff', 'color': '#111'});",
      "}"
    ),
    columnDefs = list(
      list(visible = FALSE, targets = c())
    )
  ), 
  rownames = FALSE,
  escape = FALSE,
  editable = TRUE
)


d <- d_wide


# Create crosstab for plot
      # Convert episode date to date object
      d$`Episode Date` <- as.Date(d$`Episode Date`, format = "%d-%b-%y")
      
      # Remove cases with no episode date
      d <- d %>% filter(! is.na(`Episode Date`))
      
      # Sort data by episode date
      d <- d[order(d$`Episode Date`),]
      
      # Create a crosstab
      crosstab <- d %>% group_by(`Episode Date`) %>% tally()
      
      # Rename the n vector
      names(crosstab)[ncol(crosstab)] <- "Incidence"
      
      # Compute cumulative incidence
      crosstab <- crosstab %>% mutate(`Cumulative Incidence` = cumsum(Incidence))
      
      # Create an age group vector
      crosstab <- crosstab %>% mutate(`Age Group` = rep("All ages", nrow(crosstab)))
      
      # Reorder columns
      crosstab <- crosstab %>% select(`Age Group`, everything())
      
      # Add day column
      crosstab <- crosstab %>% mutate(Day = get_days(unlist(`Episode Date`), day1 = min(crosstab$`Episode Date`)))
      
      # Create a crosstab by age group
      crosstab2 <- d %>% group_by(`Age Group`, `Episode Date`) %>% tally()
      
      # Rename the n vector
      names(crosstab2)[ncol(crosstab2)] <- "Incidence"
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Compute cumulative incidence
      #crosstab2 <- crosstab2 %>% group_by(`Age Group`) %>% mutate(`Cumulative Incidence` = cumsum(Incidence))
      
      crosstab2$`Cumulative Incidence` <- unlist(aggregate(crosstab2$Incidence, by = list(crosstab2$`Age Group`), cumsum)$x)
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Add day column
      crosstab2 <- crosstab2 %>% group_by(`Age Group`) %>% mutate(Day = get_days(unlist(`Episode Date`), day1 = min(crosstab$`Episode Date`)))
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Combine both crosstabs
      crosstab <- rbind(crosstab, crosstab2)
      
      # Ensure that the age group vector is a factor
      crosstab$`Age Group` <- factor(crosstab$`Age Group`)
      
      # Relevel the age group factor
      crosstab$`Age Group` <- relevel(crosstab$`Age Group`, ref = "All ages")


```

## Incidence plot by age group

```{r}
# Print line plot
point_size <- 0.5
element_text_size <- 12
plot_width <- 900
plot_height <- 614
ggplotly(ggplot(crosstab, aes(x = `Episode Date`, y = Incidence)) +
    geom_line(aes(color = `Age Group`), size = point_size) +
    ggtitle("Incidence by age group") +
    xlab("Date") +
    ylab("Incidence") +
    theme_minimal() +
    theme(
       plot.title = element_text(size = element_text_size),
       axis.title.x = element_text(size = element_text_size),
       axis.title.y = element_text(size = element_text_size),
       legend.text = element_text(size = element_text_size),
       legend.title = element_blank()
  ), width = plot_width, height = plot_height)
```

## Incidence plot by age group

```{r}
# Print line plot
ggplotly(ggplot(crosstab, aes(x = `Episode Date`, y = `Incidence`)) +
    geom_line(aes(color = `Age Group`), size = point_size) +
    ggtitle("Cumulative incidence by age group") +
    xlab("Date") +
    ylab("Cumulative incidence") +
    theme_minimal() +
    theme(
       plot.title = element_text(size = element_text_size),
       axis.title.x = element_text(size = element_text_size),
       axis.title.y = element_text(size = element_text_size),
       legend.text = element_text(size = element_text_size),
       legend.title = element_blank()
  ), width = plot_width, height = plot_height)
```

## Cumulative incidence plot by age group

```{r}
# Print line plot
ggplotly(ggplot(crosstab, aes(x = `Episode Date`, y = `Cumulative Incidence`)) +
    geom_line(aes(color = `Age Group`), size = point_size) +
    ggtitle("Cumulative incidence by age group") +
    xlab("Date") +
    ylab("Cumulative incidence") +
    theme_minimal() +
    theme(
       plot.title = element_text(size = element_text_size),
       axis.title.x = element_text(size = element_text_size),
       axis.title.y = element_text(size = element_text_size),
       legend.text = element_text(size = element_text_size),
       legend.title = element_blank()
  ), width = plot_width, height = plot_height)
```


```{r}

# Get data for the comparison of data snapshots
  get_data2 <- function(x) {
    if(is.null(x) | ! x %in% list.files()) {
      return(list(d = NULL, crosstab = NULL))
    } else {
      # Get data
      d <- read_xlsx(x)
      
      # Isolate date
      snapshot <- unname(sapply(x, function(x) strsplit(x, "updated ")[[1]][2]))
      snapshot <- gsub(".xlsx", "", snapshot)
      
      # Reformt the date
      snapshot <- format(as.Date(file_names), "%B %d, %Y")
      
      # Add d to list
      l <- list(d = d)
      
      # Create crosstab for plot
      # Convert episode date to date object
      d$`Episode Date` <- as.Date(d$`Episode Date`, format = "%d-%b-%y")
      
      # Remove cases with no episode date
      d <- d %>% filter(! is.na(`Episode Date`))
      
      # Sort data by episode date
      d <- d[order(d$`Episode Date`),]
      
      # Create a crosstab
      crosstab <- d %>% group_by(`Episode Date`) %>% tally()
      
      # Rename the n vector
      names(crosstab)[ncol(crosstab)] <- "Incidence"
      
      # Compute cumulative incidence
      crosstab <- crosstab %>% mutate(`Cumulative Incidence` = cumsum(Incidence))
      
      # Create an age group vector
      crosstab <- crosstab %>% mutate(`Age Group` = rep("All ages", nrow(crosstab)))
      
      # Reorder columns
      crosstab <- crosstab %>% select(`Age Group`, everything())
      
      # Add day column
      crosstab <- crosstab %>% mutate(Day = get_days(unlist(`Episode Date`), day1 = min(crosstab$`Episode Date`)), Snapshot = rep(snapshot, nrow(crosstab)))
      
      # Create a crosstab by age group
      crosstab2 <- d %>% group_by(`Age Group`, `Episode Date`) %>% tally()
      
      # Rename the n vector
      names(crosstab2)[ncol(crosstab2)] <- "Incidence"
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Compute cumulative incidence
      #crosstab2 <- crosstab2 %>% group_by(`Age Group`) %>% mutate(`Cumulative Incidence` = cumsum(Incidence))
      crosstab2$`Cumulative Incidence` <- unlist(aggregate(crosstab2$Incidence, by = list(crosstab2$`Age Group`), cumsum)$x)
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Add day column
      crosstab2 <- crosstab2 %>% group_by(`Age Group`) %>% mutate(Day = get_days(unlist(`Episode Date`), day1 = min(crosstab$`Episode Date`)), Snapshot = rep(snapshot, nrow(crosstab2)))
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Combine both crosstabs
      crosstab <- rbind(crosstab, crosstab2)
      
      # Ensure that the age group vector is a factor
      crosstab$`Age Group` <- factor(crosstab$`Age Group`)
      crosstab$Snapshot <- factor(crosstab$Snapshot)
      
      # Relevel the age group factor
      crosstab$`Age Group` <- relevel(crosstab$`Age Group`, ref = "All ages")
      
      # Add crosstab to list
      l$crosstab <- crosstab
      
      return(l)
    }
  }
  
  # Compute the day numbers from day 1
  get_days <- function(x, day1) {
    day1 <- day1 - 1
    day <- c()
    for(i in 1:length(x)) {
      day[i] <- as.numeric(x[i] - day1)
    }
    return(day)
  }


a <- get_data2("Table 13-10-0766-01 - updated 2020-04-24.xlsx")

x = "Table 13-10-0766-01 - updated 2020-04-24.xlsx"

d <- read_xlsx(x)

# Isolate date
      snapshot <- strsplit(x, "updated ")[[1]][2]
      snapshot <- gsub(".xlsx", "", snapshot)
      
      # Reformt the date
      snapshot <- format(as.Date(snapshot), "%B %d, %Y")
      
      # Add d to list
      l <- list(d = d)
      
      # Create crosstab for plot
      # Convert episode date to date object
      d$`Episode Date` <- as.Date(d$`Episode Date`, format = "%d-%b-%y")
      
      # Remove cases with no episode date
      d <- d %>% filter(! is.na(`Episode Date`))
      
      # Sort data by episode date
      d <- d[order(d$`Episode Date`),]
      
      # Create a crosstab
      crosstab <- d %>% group_by(`Episode Date`) %>% tally()
      
      # Rename the n vector
      names(crosstab)[ncol(crosstab)] <- "Incidence"
      
      # Compute cumulative incidence
      crosstab <- crosstab %>% mutate(`Cumulative Incidence` = cumsum(Incidence))
      
      # Create an age group vector
      crosstab <- crosstab %>% mutate(`Age Group` = rep("All ages", nrow(crosstab)))
      
      # Reorder columns
      crosstab <- crosstab %>% select(`Age Group`, everything())
      
      # Add day column
      crosstab <- crosstab %>% mutate(Day = get_days(unlist(`Episode Date`), day1 = min(crosstab$`Episode Date`)), Snapshot = rep(snapshot, nrow(crosstab)))
      
      # Create a crosstab by age group
      crosstab2 <- d %>% group_by(`Age Group`, `Episode Date`) %>% tally()
      
      # Rename the n vector
      names(crosstab2)[ncol(crosstab2)] <- "Incidence"
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Compute cumulative incidence
      #crosstab2 <- crosstab2 %>% group_by(`Age Group`) %>% mutate(`Cumulative Incidence` = cumsum(Incidence))
      crosstab2$`Cumulative Incidence` <- unlist(aggregate(crosstab2$Incidence, by = list(crosstab2$`Age Group`), cumsum)$x)
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Add day column
      crosstab2 <- crosstab2 %>% group_by(`Age Group`) %>% mutate(Day = get_days(unlist(`Episode Date`), day1 = min(crosstab$`Episode Date`)), Snapshot = rep(snapshot, nrow(crosstab2)))
      
      # Restructure as tibble
      crosstab2 <- as_tibble(crosstab2)
      
      # Combine both crosstabs
      crosstab <- rbind(crosstab, crosstab2)
      
      # Ensure that the age group vector is a factor
      crosstab$`Age Group` <- factor(crosstab$`Age Group`)
      crosstab$Snapshot <- factor(crosstab$Snapshot)
      
      # Relevel the age group factor
      crosstab$`Age Group` <- relevel(crosstab$`Age Group`, ref = "All ages")
      
      # Add crosstab to list
      l$crosstab <- crosstab
      
      return(l)
```